A new way to speedup in global optimization problems

### 2019-03-06 2019/NUMTA_Barkalov/Abstract/Barkalov.tex
Russian text

@@

\title{Title of Your Presentation}
\author{Roman Strongin, Konstantin Barkalov, Semen Bevzuk}

@@

\institute{University of Nizhni Novgorod, Nizhni Novgorod, Russia\\
\email{president@unn.ru, barkalov@vmk.unn.ru}}

@@

\textbf{Keywords.} Global optimization; keyword2; keyword3.

@@

The goal of the Conference is to create a multidisciplinary round
table for an open discussion on numerical modeling nature by using
traditional and emerging computational paradigms. The Conference
will discuss all aspects of numerical computations and modeling from
foundations and philosophy to advanced numerical techniques. New
technological challenges and fundamental ideas from theoretical
computer science, linguistic, logic, set theory, and philosophy will
meet requirements and new fresh applications from physics,
chemistry, biology, and economy.

Researchers from both theoretical and applied sciences are very
welcome to submit their papers and to use this excellent possibility
to exchange ideas with leading scientists from different research
fields. Papers discussing new computational paradigms, relations
with foundations of mathematics, and their impact on natural
sciences are particularly solicited. A special attention will be
also dedicated to numerical optimization and different issues
related to theory and practice of the usage of infinities and
infinitesimals in numerical computations.

@@

\bibitem{Strongin18} Barkalov K., Strongin R. (2018) Solving a set of global optimization problems by the parallel technique with uniform convergence. \emph{J. Glob. Optim.}, Vol.~71(1), pp.~21--36.

@@

>>>>

\usepackage[utf8x]{inputenc}
\usepackage[english,russian]{babel}
\usepackage{cmap}

@@

\title{A new way to speedup in global optimization problems}
\author{Roman Strongin, Konstantin Barkalov, Semen Bevzuk}

@@

\institute{University of Nizhni Novgorod, Nizhni Novgorod, Russia\\
\email{barkalov@vmk.unn.ru}}

@@

\textbf{Keywords.} Global optimization; multiextremal problems; Lipschitz constant.

@@

В работе рассматриваются задачи глобальной оптимизации c black-box objective function, удовлетворяющей условию Липшица. Важной проблемой в задачах данного класса является адекватная оценка априори неизвестной константы Липшица на основе результатов проведенных search trials. В монографии \cite{Book} предложены различные подходы к учету как глобальных, так и локальных свойств objective function. В частности, показали свою перспективность  алгоритмы с использованием локальных оценок константы Липшица \cite{Sergeyev10, Sergeyev16}.

В данной статье предлагается новый способ учета локальных свойств objective function. Предложенный способ основан на использовании в информационно-статистическом алгоритме глобального поиска \cite{Book} одновременно двух оценок константы Липшица: нижней и верхней. Правила нового алгоритма обеспечивают автоматическое переключение между указанными двумя оценками в зависимости от поведения оптимизируемой функции и от стадии поиска.

Предложенный новый алгоритм гарантирует сходимость к global minimizer. Результаты решения серии тестовых задач разной размерности подтверждают как наличие глобальной сходимости  нового алгоритма, так и его ускорение по сравнению с исходным алгоритмом.

@@

@@

%\bibitem{Strongin18} Barkalov K., Strongin R. (2018) Solving a set of global optimization problems by the parallel technique with uniform convergence. \emph{J. Glob. Optim.}, Vol.~71(1), pp.~21--36.
